---
title: "p8105_hw1_mfc2162"
output: github_document
---

```{r}
library(tidyverse)
```

## Problem 1

#### Create dataframe

```{r}
df1 = tibble(
  norm_samp = rnorm(8),
  vec_logical = norm_samp > 0,
  vec_char = c("There's", "always", "money", "in", "the", "banana", "stand",
               "!"),
  vec_factor = factor(c("medium", "low", "high", "medium", "low", "low",
                        "medium", "medium"))
)

df1
```


#### Take the mean of each variable

```{r}
mean_samp = mean(pull(df1, norm_samp))
mean_logical = mean(pull(df1, vec_logical))
mean_char = mean(pull(df1, vec_char))
mean_factor = mean(pull(df1, vec_factor))

mean_samp
mean_logical
mean_char
mean_factor
```

Taking the mean of the random sample variable and the logical variable worked.
Taking the mean of the character variable and the factor variable did not work.

#### Apply as.numeric to variables

```{r}
logical_num = as.numeric(pull(df1, vec_logical))
char_num = as.numeric(pull(df1, vec_char))
factor_num = as.numeric(pull(df1, vec_factor))
```

It works for the logical and factor variables, but not for the character
variable. Applying as.numeric to the logical variable just changes the TRUE's to
1's and the FALSE's to 0's. Applying as.numeric to the factor variable assigns a
number to each level. In this case, it assigned 3 to "medium", 2 to "low", and 1
to "high". Applying as.numeric to a character vector only works if the characters
are numbers, such as c("3", "6", "8", "3", "7"). In this case, the characters
were not numbers, so it didn't work.


#### Conversion and multiplication

```{r}
logical_num = as.numeric(pull(df1, vec_logical))
product_num = logical_num * pull(df1, norm_samp)
product_num

logical_fac = as.factor(pull(df1, vec_logical))
product_fac = logical_fac * pull(df1, norm_samp)
product_fac

logical_fac_to_num = as.numeric(logical_fac)
product_fac_to_num = logical_fac_to_num * pull(df1, norm_samp)
product_fac_to_num
```

## Problem 2

#### Create data frame

```{r}
df2 = tibble(
  x = rnorm(500),
  y = rnorm(500),
  log_vec = x + y > 1,
  num_vec = as.numeric(log_vec),
  fac_vec = as.factor(log_vec)
)

df2
```

#### Data description

The number of rows in the dataset is `r nrow(df2)` and the number of columns is
`r ncol(df2)`.

The mean of x is `r mean(pull(df2, x))`.
The median of x is `r median(pull(df2, x))`.
The standard deviation of x is `r sd(pull(df2, x))`.

The proportion of cases for which x + y > 1 is `r sum(pull(df2, log_vec))/500`.

#### Scatterplots


Color by logical variable
```{r}
scatter1 = ggplot(df2, aes(x = x, y = y, color=log_vec)) + geom_point()
scatter1
```


Color by numeric variable
```{r}
ggplot(df2, aes(x = x, y = y, color=num_vec)) + geom_point()
```

The color scale is continuous because numeric variables are continuous. However,
all of the points are either light blue or dark blue because they are all 1 or
0.


Color by factor variable
```{r}
ggplot(df2, aes(x = x, y = y, color=fac_vec)) + geom_point()
```

The color scale is discrete because factor variables are made up of discrete
levels. Points are red if FALSE and blue if TRUE.

#### Export first scatterplot

```{r}
ggsave("HW1_scatterplot.pdf", plot = scatter1, height = 4, width = 6)
```